---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

```{r cars}
summary(cars)
```

```{r pressure, echo=FALSE}
plot(pressure)
```

```{r}
library(ggplot2)
library(dplyr)

```


## Including Code
```{r}
library(readr)
clean_data <- read.csv("~/GitHub/ppol8102-Paul/codebook/clean_data.csv")
View(clean_data)
```

##Inspect the data
```{r}
head(clean_data)
```

```{r}
str(clean_data)
```

```{r}
summary(clean_data)
```

##Description
This data contains variables on Preventive care utilization from NHANES 2017 - 2018 dataset. The variables include age, gender, education, income,marital status,race, general health condition,mental health,oral health and health insurance.

## Variable Description
Dependent variables: General Health. This is an ordinal categorical variable
                     Mental Health. This is Binary
                     Oral Health: This is a categorical variable
Independent variable:  Race. This is a categorical variable ( Black or White)
Control variables: Age, gender, income, education, marital status, health insurance. These are categorical                      variables.

You can include R code in the document as follows:


## Converting Race to binary
```{r}
clean_data$race <- as.factor(clean_data$race)
levels(clean_data$race)
```
```{r}
clean_data$health <- as.factor(clean_data$health)
levels(clean_data$health)
```
```{r}
str(clean_data$health)
```

#Convert health from factor to numeric
```{r}
clean_data$health <- as.numeric(as.character(clean_data$health, exclude = NULL))
str(clean_data$health)
```

##Summary statistics for general health by race
```{r}
clean_data %>%
  group_by(race) %>%
  summarise(
    Mean_health = mean(health, na.rm = TRUE),
    SD_health = sd(health, na.rm = TRUE),
    N = n()
  )
```

# Check for missing values in the health variable
```{r}
sum(is.na(clean_data$health))
sum(!is.finite(clean_data$health))
```

##Removing rows with missing values and imputing missing values
```{r}
clean_data <- clean_data %>% filter(is.finite(health))
clean_data$health[is.na(clean_data$health)] <- mean(clean_data$health, na.rm = TRUE)
```

##Boxplot of Health by Race
```{r}
ggplot(clean_data, aes(x = race, y = health)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "Comparison of General Health by Race",
    x = "Race",
    y = "Health (1 = Poor, 5 = Excellent)"
  )
```




```{r}
clean_data_oral <-factor(clean_data$oral, levels = c("6 months or less", "more than 6 months"),
                                     labels = c("6 months or less", "more than 6 months"), exclude = NULL)
```

```{r}
clean_data_oral <- ifelse(clean_data_oral== "6 months or less", 0, 1)
table(clean_data_oral, useNA = "always")
sum(is.na(clean_data$oral))
str(clean_data_oral)
```



You can also embed plots, for example:


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
