---
title: "Untitled"
output: github_document
author: sarah paul
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
data <- read.csv("codebook/clean_data.csv")
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(broom)
library(boot)
```

```{r}
#install.packages("rstanarm")
library("rstanarm")
```

```{r}
data_oral <- factor(data$oral, levels = c("6 months or less", "more than 6 months"), 
                    labels =c(0,1), exclude = NULL)
data$oral <- as.numeric(data$oral == "6 months or less")
print(data)
```

##fitting a regression
```{r}
model_1 <- stan_glm(oral~race, data=data)
print(model_1)
```
```{r}
print(model_1, digits = 2)
```

```{r}
race <- as.numeric(data$race == "Blacks")
```



```{r}
plot(race,data$oral, main = "Data and fitted regression line")
a_hat <-coef(model_1)
b_hat <-coef(model_1)
abline(a_hat, b_hat)

race_bar <- mean(race)
text(race_bar, a_hat+ b_hat* race_bar,
     paste("race =", round(a_hat, 2), "+" , round(b_hat, 2), "* race"), adj= 0)
```

##compute R2
```{r}
R2 <- 1 - sigma(model_1)^2 / sd(data$oral, na.rm=T)^2
print(R2)
```

## The R2 explains a small portion of the variance