---
title: "Untitled"
output: github_document
author: sarah paul
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
data <- read.csv("C:/Users/sarad/OneDrive/Desktop/Studio 2/clean_data.csv")
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(broom)
library(boot)
library("rstanarm")
library(reshape2)

```

```{r}
#install.packages("mice")
#install.packages("bayesplot")
library(mice)
library(bayesplot)
```


```{r}
data_oral <- factor(data$oral, levels = c("6 months or less", "more than 6 months"),
                    labels = c(0,1), exclude = NULL)
data$oral <- as.numeric(data$oral == "6 months or less")
print(data)
```

##predictive mean matching for missing variables
```{r}
imputed_data <- mice(data , method = "pmm", m = 5 )
complete_data <- complete(imputed_data)
```

```{r}
model_1 <- stan_glm(oral ~ race + gender + age +educ + income + marital_status + insurance,
                    data = data , 
                    family = gaussian(),
                    weights = weight ,
                    chains = 4, iter = 2000, seed = 1234)
summary(model_1)
```

## Interpretation of result
Race does not significantly predict oral health disparities as the estimate and standard deviation are 0.0. Lack of health insurance is associated with worse oral health ( SD = 0.2). Higher education is associated with better oral health. Older age groups had slightly lower oral health score than younger age group.

```{r}
fit_1 <- stan_glm(oral ~ race, data = data)
print(fit_1)
```

```{r}
fit <- stan_glm(oral ~ 1, data = data)
sims <- as.matrix(fit)
n_sims <- nrow(sims)
n <- length(data$y)
oral_rep <- array(NA, c(n_sims, n))
for (s in 1:n_sims) {oral_rep[s,] <- rnorm(n, sims[s,1], sims[s,2])
}
```

```{r}
length(na.omit(oral_rep[s, ]))
print(class(oral_rep[s, ]))  # Check if it's numeric
print(str(oral_rep[s, ])) 
```


```{r}
oral_rep_df <- as.data.frame(oral_rep)  # Convert matrix to data frame
oral_rep_df$sim_id <- 1:nrow(oral_rep_df)
oral_rep_long <- melt(oral_rep_df, id.vars = "sim_id", variable.name = "Observation", value.name = "Simulated_Oral")
# Check if oral_rep_long was created properly
head(oral_rep_long)
```

```{r}
plot(model_1$residuals, model_1$y)
```



```{r}
 ggplot() +
  geom_density(aes(x = oral_rep_long$value), fill = "blue", alpha = 0.5) +
  geom_density(aes(x = data$oral), color = "red", size = 1) +
  labs(title = "Comparison of Simulated vs Observed Data",
       x = "Oral Values",
       y = "Density") +
  theme_minimal()
```


```{r}
sims <- as.matrix(fit_1)
predicted <- predict(fit_1)
resid <- data$oral - predicted
print(resid)
plot(sims)
```



```{r}
colnames(data)
```


```{r}
model <- lm(oral ~ race + educ + age + gender + marital_status + income + insurance, 
            data = data, weights =  weight, na.action = na.exclude)
summary(model)
```
## Education and income significantly improve oral health, with higher education being strongly associated with better outcomes. Age, race, gender, marital status, and insurance all play a role in oral health. Individuals aged 20-39 experience the sharpest decline, while white individuals have better oral health than non-White individuals. Males have slightly worse oral health than females, and those without insurance have worse oral health, but the effect is not as significant.





