---    
title: "Untitled"
output: github_document
author: "Sarah Paul"
---
    
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

##Install the Bootstrap package
```{r}
#install.packages("boot")
install.packages("rstanarm")
library("boot")

```

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(broom)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
clean_data <- read.csv("~/GitHub/ppol8102-Paul/codebook/clean_data.csv")
View(clean_data)
```

##Dependent variable
Conceptualization:
General Health Condition: This is the overall state of physical, mental and emotional well being of an individual.

Operationalization: This is measured by asking respondents how they will rate their/spouse health in general from from a scale of Excellent to poor.

```{r}
str(clean_data$health)
```

##convert the dependent variable from categorical to numeric
```{r}
clean_data$health <- as.numeric(factor(clean_data$health, 
    levels = c("Excellent", "Very Good", "Good", "Fair", "Poor"), 
    labels = c(5, 4, 3, 2, 1), exclude = NULL))
str(clean_data$health)
```


```{r}
mean_clean_data_health <- mean(clean_data$health, na.rm=T)
```

```{r}
health <- clean_data$health
```

## Simulating the data using bootstrapping
```{r}
n <-nrow(clean_data)
boot<- sample(n, replace = TRUE)
health_boot <- health[boot]
ratio_boot <- mean(health_boot, na.rm = T)
```


```{r}
boot_ratio <- function(data){
n <- nrow(data)
boot <- sample(1:n, replace=TRUE)
health_boot <- data$health[boot]
mean(health_boot, na.rm = T)}

```

```{r}
n_sims <- 5000
output <- replicate(n_sims, boot_ratio(data=clean_data))
hist(output, col = "pink")
print(sd(output))

```

##Confidence Interval
```{r}
 (quantile(output, probs = c(0.025, 0.975)))
```

