---
title: "Codebook"
output: github_document
---     Author: Sarah Paul

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Open the file

```{r}
library(readr)
eclsk <- read_csv("~/GitHub/ppol8102-Paul/codebook/eclsk.csv")
View(eclsk)
```


```{r}
 summary(eclsk)

```
## Description
This data contains information on math and reading scores of students, Parent education whether high school, Bachelors of Graduate school, income category, age, sex.

##Variable description

Dependent variable : read score. This is a continous variable 
Independent variable: Parent education. This is a categorical variable. Where High school or less = A.  College = B, Graduate = C
Control variable: Imcome. This is a categorical variable andgrouped into Low, middle or high
```{r}

names(eclsk)
```
## plot a graph
```{r}
boxplot(read~parent.ed, data= eclsk)
```
##create a density plot
```{r}
plot(density(eclsk$read))
```


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
